paper

### 2019-01-11 Report/paper.tex
push paper and figures

\begin{abstract}
    In this report we investigate ways to automatically analyze ICLS and
CSCL papers to gain an understanding of the community formed by
conference participants and of the research being conducted within the
community. After describing how to extract data from the base dataset we
derive insights into the community using network analysis, natural
language processing and text mining. We discover patterns of
collaboration between participants, their migration and the global
spread of the community. We extract keywords and find clusters of
referenced papers and documents within the conferences.
\end{abstract}
    \hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The starting point of the project are the proceedings of the
International Conference of the Learning Sciences (ICLS), respectively
the International Conference on Computer-Supported Collaborative
Learning (CSCL). The two conferences are organized by the International
Society of the Learning Sciences (ISLS) and held in biennial alternation
with each other. At the time of writing the dataset contains 4 years of
the proceedings, from 2015 to 2018. While ICLS covers the entire field
of the learning sciences, CSCL focuses on learning through collaboration
with the help of communication technologies \cite{ICLS}. To understand
the community formed by the participants of the conferences as well as
to understand the research being conducted within the community, we
explore methods to bring insight into the following aspects:

@@

    The provided raw dataset consists of papers in \emph{pdf} format, as
well as associated metadata in \emph{xml} format. To extract information
from the dataset we first converted the data to a format that is better
adapted to the task of data analysis and more compatible with a variety
of external libraries. We stored all the resulting data in \emph{csv}
format, as csv is a human readable format and can easily be processed by
all the data analysis tools we employ.

@@

We extracted the metadata from the xml files using the \emph{lxmltree}
python library. We associated the contents of each xml tag to a column
in a csv file. As the xml files do not exactly follow the xml standard,
we handled errors by iteratively parsing the xml tree. For some files,
the section containing the keywords is malformed and can not be
extracted by the parser. This was remedied using regular expressions to
match the section containing the keywords. To each file we associated
the name of the source file and the order in which an author name occurs
in the xml file. Each author is associated to one line in the xml file.
From the associated citation string we add the first part of the
citation containing the author names, year, and title. This is later
used to identify the paper when it is referenced in an other paper. Then
we add a shortened version of author names (i.e.~West, R.) by
considering the author order. Note that it happens that the short name
extracted from the citation and the corresponding long name (Robert
West) in the metadata do not match up. Thus the author order listed in
the citation string does not always correspond to the author order
listed in the metadata. After extraction, we cleaned the data by
unifying names, as slight spelling differences exist within the same
author name (i.e.~Robert West, Robert A. West). When unifying the names
we needed to make sure not to overmerge by labeling two different
authors as having the same name. First, we used a strict condition to
catch small differences in names, such as an extra middle name. This is
done by splitting up names into substrings based on whitespaces and
commas. For two names in the dataset we checked whether the intersection
of their sets of substrings contained at least 2 names of length at
least 3. If the condition was met we considered these two names to be
identical, and added the mapping from one name to the other to a
dictionary. Once all tuples of names (s.t. the first name appears before
the second one in order) were checked and added to the dictionary we
unified the names. Using this method we reduced the number of authors in
the dataset from 1951 down to 1879. Further, to merge names with
misspellings or people who use nicknames (Christian vs Chris) and to
avoid overmerging we considered only names of authors that have similar
collaboration patterns. To do this we built a graph based on
co-authorship relations and checked names of authors that were in the
same neighborhood of a node on the graph (see methodology for further
details). Then, we used the \emph{difflib} library to check the
similarity of two strings and considered a name to belong to the same
person if the similarity measure reached the threshold of 0.8. Again,
names considered to belong to the same person were added to the mapping
dictionary which was then applied to all names. This method again
reduced the number of authors in the dataset from 1879 to 1851.

@@

Now we consider the section before the abstract, which contains the
authors names, their email addresses and the institutions they are
affiliated with. Affiliation with institution can give us a lot of
insight where people are working and gives use a way estimate where they
are from. To extract the exact affiliation of an authors with an
institution two approaches were used.

@@

The most robust way to get a persons institution affiliation is to use
their email address. Email addresses can easily be extracted from the
header section using a regular expression. Then, the part of the email
containing the university domain can be extracted using an other regular
expression. After this we used a domain-university mapping
\cite{mapping} to map domains to institutions. The additional advantage
of this method is that using the additional metadata associated to each
university in the mapping, we also access information about the country
of the university.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_11_0.png}\end{center}
        \caption{Different formats used to reference institution}
        \label{figA}
    \end{figure}

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_12_0.png}\end{center}
        \caption{Different formats used to reference institution}
        \label{figB}
    \end{figure}

@@

The header section itself contains information on the institution, but
different ways to list authors are accepted, notably two ways as can be
seen in Figure \ref{figA} and Figure \ref{figB}. Hence, we first
detected the structure of the header section before extracting the
association. We tagged the substring present in this section using a
combination of \emph{named entity detection}, which automatically
classifies strings as belonging to some pre-defined category, using the
implementation provided by \emph{spacy}\cite{spacy2}, and simple string
comparison matching of the names of authors and institutions we
extracted from the metadata respectively from Approach 1 and tagging
sections containing ``emails'' using regex , to detect where in the
header the email is. Then, taking the tagged section, we detected which
subsections corresponded to which author-affiliation listing type, and
extracted the author email - institution mapping accordingly.

@@

As there are a variety of way in which the name of an institution could
be written --consider EPFL, EPF Lausanne-- crawling the english, french
and german version of Wikipedia (not using the API, or the API wrapper,
as this does not give access to the desired information) was a quick and
accurate way to get the location. The advantage of sending requests over
using the API is due to the website redirecting to the relevant article
on the institution even when misspellings are present or multiple ways
to spell the university exist. Requests were sent using \emph{urllib3},
and \emph{beautifulsoup} was used to parse them get the coordinates
present in the article. Regretfully not all institutions, even well
known ones such as Tokyo University, have their coordinates listed on
Wikipedia. To get the remaining locations we used geopy, which is a
python client for geocoding web services. These results however are more
error prone as geopy will always return a result. It is thus recommended
to verify the correctness of the results. To get the country we then
again used geopy. The resulting data containing paper identifier, the
authors email, its associated domain, order in which the author was
cited, the institution, country, longitude and latitude were saved to a
csv file. We used the paper identifier and author order to associate
this data with the metadata to link information about authors with the
papers they authored.

@@

To get insight into collaborations within the community we built graphs
based on co-authorship of papers. To build the graph structure and
visualize it the python package \emph{networkx} was used. When
constructing the graph, we associate qualitative features --such as the
name and institution of the author that is represented by a node-- to
the nodes and edges of the graph. We also associate the number of
collaborations between different parties to the graphs as the weight of
the edge.

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_46_0.png}\end{center}
        \caption{One of the connected components of the co-authorship graph}
        \label{fig_plot}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_48_0.png}\end{center}
        \caption{Largest Component}
        \label{author_collab_large}
    \end{figure}

    Focusing in on the authors is not the only way to interpret the
co-authorship graph. One should also consider the overall structure of
the graph. As the co-authorship graph is not connected, and contains one
large main component with 1141 authors and 208 very small components
with less than 25 authors in them we focused on analyzing the structure
of the main component of the graph. First we aimed to detect
sub-communities within the graph. Modularity is one measure used to find
such communities. It measures the connection strength within a community
relative to the strength of outgoing connections. Networks with high
modularity have dense connections between the nodes within the
subcommunites but sparse connections between nodes in different
subcommunites. A non-parametric algorithm for community detection is the
Louvain Method \cite{blondel2008fast}. We used the implementation of
Thomas Aynaud \cite{louvain}. This gives a good partitioning of the
graph into 26 communities of median size 42.84. This is quite a large
number of communities, so we considered an alternative way to yield
bigger partitions. For this we looked at normalized cuts, which
partition the nodes into groups with many within-group connections and
relatively few between-group connections. The advantage of this method
is that we can choose how many cuts to make, and we are thus able to
define the size of subcommunities. Finding normalized cuts in a graph
can be done using \emph{sklearns} spectral clustering algorithm by
passing it the adjacency matrix of the graph. With this method we are
able to neatly separate the main component graph in three large
communities, as can be seen in figure \ref{normalized}

@@

To get a better overview of how co-authorship happens across countries
and institutions, we built graphs where every node represented a
country, respectively an institution. Two countries or institutions were
linked if two authors from these two institutions collaborated on a
paper together. The country graph is simple enough that it does not have
to be pruned. The institution graph on the other hand is again not
comprehensible without further processing. Thus, similar to the first
approach on pruning presented, we only considered institutions who
collaborated in at least two conferences. This the main component of
this graph is shown in figure \ref{unigraph}.

@@

Graphs were visualised with the built in plot functions and graph
layouts of networkx. The force directed layout was used as it clusters
together nodes with strong connections. As this layout is produced by
iterations and based on random initial conditions, we set a fix random
seed to be able to reproduce the positioning. The size of each node was
changed based on the position of the node in the graph. Central nodes
with a lot of connections should be larger, while nodes with fewer
connections should be smaller. To achieve this we used the measures of
degree centrality for graphs with few nodes and edges and betweenness
centrality for larger graphs. Degree centrality measures the fraction of
degrees (number of connections) a node has with respect to the total
number of connections in the graph. Betweenness centrality of a node is
the sum of the fraction of all-pairs shortest paths that pass through it
\cite{borgatti2014social}. We found that degree centrality is more
appropriate for smaller graphs as their structure is still easy to
interpret. As by construction betweenness takes into consideration the
degrees of it neighboring nodes, it better structures the nodes large
graphs into central and less central nodes. The width of the edge is
based on the weight of the edge, which in turn reflects the number of
collaboration between two nodes.

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_59_0.png}\end{center}
        \caption{Number of Contributiors by coutry, excluding the US}
        \label{figrep}
    \end{figure}

@@

\textbf{International Collaboration:} The U.S. again stands out as a
major player, with most collaborations happening between the U.S. and
other countries. Out of 851 papers, 153 were international
collaborations. Out of these 153, 108 involved at least one U.S.
participant. Excluding US participation we find the collaboration graph
presented in figure \ref{figcollab}. We note that some countries, such
as Japan, are not present in this graph. This means that all
international collaboration of these countries, in this case japanese
authors, also involves American authors. Additionally we observed that
many Asian nations seem to collaborate almost exclusively with each
other or Canada.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_61_0.png}\end{center}
        \caption{Collaborative papers between countries excluding US}
        \label{figcollab}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_64_0.png}\end{center}
        \caption{Subcommunities by paricipation in CSCL. Blue: no participation in main compnents of CSCL. Red and orange each mark participation in one of the two main components of CSCL}
        \label{CSCL_components}
    \end{figure}

@@

The fact that normalized graph cuts will split the community into three
promted further investigation. As will be discussed in the next section,
the graph of authors that participanted in CSCL yields two large
connected components. By coloring the authors contained in the two CSCL
components different colors, and authors in neither a third color, we
find a similar partitioning of the graph, as can be observed in figure
\ref{CSCL_components}. This is somewhat trivial, as the graph is built
based on collaboration in these conferences. However, the analysis of
nationalities present in the graph yielded an interesting results. All
Asian participants from nations other than Singapore are contained in
the same cluster. We also note that one of the CSCL clusters is closer
to the group of ICLS-only participants than the other. Additionaly,
through normalized splits, we find that the section of the normalized
cut containing ICLS only authors is overwhelmingly american.

Considering the wordclouds in figure \ref{c1}, \ref{c2}, \ref{c3} of the
most common keywords in each of the large splits, we find a first
indication that the splits also reflect some thematic differences in
research.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_66_0.png}\end{center}
        \caption{Keywords of the blue Community}
        \label{c1}
    \end{figure}

@@

In our analysis, we could find many structural differences between the
two conferences. Firstly, ICLS is much larger and american dominated
than CSCL. We also observed differences in the most cited papers and the
average number of authors of a paper. By constructing the graph based on
collaboration of authors in each conference, we find that while ICLS
contains one giant component, CSCL contains two large main components.
All these results are however to be taken with a grain of salt, as for
each conference we only have data for two years, and the year 2018 is a
marked outlier. Additionaly, we are unable to find a strong difference
between the metrics of the graphs, that is diameter, density, clustering
coefficient and average path length.

@@

Using the data we have of participants over the years, and knowing their
affiliation, we can now find members that change institutions over the
years. Over the years of recording we find 14 participants that have
changed countries and 47 that have moved institutions. We can see that
most movement occured within the US respectively Europe, see figure
\ref{fig_move}.

@@

We can observe a growth of the number of participants over the years,
and by building co-authorship graphs for each year of the conference, we
observe that the density and clustering coefficient of the graph
indicate a steady decrease, while the diameter and average shortest path
length tend to increase. This hints that the community is attracting new
member that do not collaborate with the established group of
participants. But, as the years 2018 shows a massive increase of
participation, in number of participants and number of institutions
represented, as well as a large decrease in density of the graph --which
is reduced by more than half from the previous ICLS conference-- it is
impossible to draw any statisically significant conclusions.

@@

    Since the number of documents is increasing over the years, we also
wanted to investigate more if the relevance of all the documents
published is increasing too. Using Pagerank algorithm and the In-degree
of each nodes to get the most influencial documents referenced of the
exploratory citations graph that we constructed before. We filtered the
first 100 documents and then contructed a histogram of the counts by
year. We obtain the figure \ref{fig7}:

@@

    We also noticed from the first graph that some documents were more
likely to cited together. We can say that the more two documents are
likely to be cited together the more it is probable that both documents
are about the same subject. From the table \ref{fig8} we can also notice
that some papers are more likely to cite distinct publications of the
same author, which is logic in a sense, because if some authors always
treat about the same subject then they are experts in that field and a
lot of their publications are relevant and can be cited together.

@@

    Using this graph, we also want to extract if there are some clusters of
papers cited together. As we explained before, we run strongly connected
algorithm to check if some cited documents are more likely to be cited
together, and this will give us clusters of cited documents that are
more likely to be about the same subjects. We can visualise one cluster,
the most important one in figure \ref{fig12}.

@@

In summary, we were able to extract how the dominant some institutions
and countries are within the conference. Some initiall collaboration
patterns between countries, instututions and authors were extracted.
Subcommunities found within the network of co-authorship reflect the
division between conferences and within the conference. Using the
analysis of the affiliation over the years we followed the migration of
participants accross institutions. Natural language preprocessing and
metadata parsing enabled the extraction of trends of keywords over the
years

Overall, due to lack of data, we were often unable find non trivial
results; This lack of data is exasperated by the fact that the
conference of 2018 presents as an outlier in the dataset. Additionally,
taking document clustering for instance as an example, network analysis
of references were more accurate than(NLP) regarding documents and
references clustering.

>>>>

\begin{abstract}
    In this report we investigate ways to automatically analyze ICLS and
CSCL papers to gain an understanding of the community formed by
conference participants and of the research being conducted within the
community. After describing how to extract data from the base dataset we
derive insights into the community using network analysis, natural
language processing and text mining. We discovered patterns of
collaboration between participants, their migration and the global
spread of the community. We extracted keywords and found clusters of
referenced papers and documents within the conferences.
\end{abstract}
    \hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The starting point of the project are the proceedings of the
International Conference of the Learning Sciences (ICLS), respectively
the International Conference on Computer-Supported Collaborative
Learning (CSCL). The two conferences are organized by the International
Society of the Learning Sciences (ISLS) and held in biennial alternation
with each other. At the time of writing the dataset contains 4 years of
the proceedings, from 2015 to 2018. While ICLS covers the entire field
of the learning sciences, CSCL focuses on learning through collaboration
with the help of communication technologies \cite{ICLS}. To understand
the community formed by the participants of the conferences as well as
to understand the research being conducted within the community, we
explored methods to bring insight into the following aspects:

@@

    The provided raw dataset consists of papers in \emph{pdf} format, as
well as associated metadata in \emph{xml} format. To extract information
from the dataset we first converted the data to a format that is better
adapted to the task of data analysis and more compatible with a variety
of external libraries. We stored all the resulting data in \emph{csv}
format, as csv is a human readable format and can easily be processed by
all the data analysis tools we used.

@@

We extracted the metadata from the xml files using the \emph{lxmltree}
python library. We associated the contents of each xml tag to a column
in a csv file. As the xml files do not exactly follow the xml standard,
we handled errors by iteratively parsing the xml tree. For some files,
the section containing the keywords is malformed and can not be
extracted by the parser. This was remedied using regular expressions to
match the section containing the keywords. To each file we associated
the name of the source file and the order in which an author name occurs
in the xml file. Each author is associated to one line in the xml file.
From the associated citation string we added the first part of the
citation containing the author names, year, and title. This is later
used to identify the paper when it is referenced in an other paper. Then
we added a shortened version of author names (i.e.~West, R.) by
considering the author order. Note that it happens that the short name
extracted from the citation and the corresponding long name (Robert
West) in the metadata do not match up. Thus the author order listed in
the citation string does not always correspond to the author order
listed in the metadata. After extraction, we cleaned the data by
unifying names, as slight spelling differences exist within the same
author name (i.e.~Robert West, Robert A. West). When unifying the names
we needed to make sure not to overmerge by labeling two different
authors as having the same name. First, we used a strict condition to
catch small differences in names, such as an extra middle name. This is
done by splitting up names into substrings based on whitespaces and
commas. For two names in the dataset we checked whether the intersection
of their sets of substrings contained at least 2 names of length at
least 3. If the condition was met we considered these two names to be
identical, and added the mapping from one name to the other to a
dictionary. Once all tuples of names (s.t. the first name appears before
the second one in order) were checked and added to the dictionary we
unified the names. Using this method we reduced the number of authors in
the dataset from 1951 down to 1879. Further, to merge names with
misspellings or people who use nicknames (Christian vs Chris) and to
avoid overmerging we considered only names of authors that have similar
collaboration patterns. To do this we built a graph based on
co-authorship relations and checked names of authors that were in the
same neighborhood of a node on the graph (see methodology for further
details). Then, we used the \emph{difflib} library to check the
similarity of two strings and considered a name to belong to the same
person if the similarity measure reached the threshold of 0.8. Again,
names considered to belong to the same person were added to the mapping
dictionary which was then applied to all names. This method again
reduced the number of authors in the dataset from 1879 to 1851.

@@

Now we consider the section before the abstract, which contains the
authors names, their email addresses and the institutions they are
affiliated with. Affiliation with institution can give us a lot of
insight where people are working and gives use a way estimate where they
are from. To extract the exact affiliation of an author with an
institution two approaches were used.

@@

The most robust way to get a person's institution affiliation is to use
their email address. Email addresses can easily be extracted from the
header section using a regular expression. Then, the part of the email
address containing the university domain can be extracted using an other
regular expression. After this we used a domain-university mapping
\cite{mapping} to map domains to institutions. The additional advantage
of this method is that using the additional metadata associated to each
university in the mapping, we also access information about the country
of the university.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_11_0.png}\end{center}
        \caption{First format used to refernce institution}
        \label{figA}
    \end{figure}

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_12_0.png}\end{center}
        \caption{Second format used to refernce institution}
        \label{figB}
    \end{figure}

@@

The header section itself contains information on the institution, but
different ways to list authors are accepted, notably two ways as can be
seen in Figure \ref{figA} and Figure \ref{figB}. Hence, we first
detected the structure of the header section before extracting the
association.

We tagged the substring present in this section using a combination of
two methods. Firstly, \emph{named entity detection}, which automatically
classifies strings as belonging to some pre-defined category, using the
implementation provided by \emph{spacy}\cite{spacy2}. This taggs
categorized strings as organisation or a person. Secondly, we used
simple string comparison matching of the names of authors and
institutions to determine if a string is an author name (person) or an
institution (organization). We extracted the data needed for this from
the metadata respectively from Approach 1. To detect sections that are
email adresses, we used regular expression. We assigned these strings
the caregory ``Email''. Then, taking the tagged section, we detected
which subsections corresponded to which author-affiliation listing type,
and extracted the authors email address - institution mapping
accordingly.

@@

As there are a variety of way in which the name of an institution can be
written --consider EPFL, EPF Lausanne-- crawling the english, french and
german version of Wikipedia (not using the API, or the API wrapper, as
this does not give access to the desired information) was a quick and
accurate way to get the location. The advantage of sending requests over
using the API is due to the website redirecting to the relevant article
on the institution even when misspellings are present or multiple ways
to spell the university exist. Requests were sent using \emph{urllib3},
and \emph{beautifulsoup} was used to parse them and get the coordinates
present in the article. Regretfully not all institutions, even well
known ones such as Tokyo University, have their coordinates listed on
Wikipedia. To get the remaining locations we used geopy, which is a
python client for geocoding web services. These results however are more
error prone as geopy will always return a result. It is thus recommended
to verify the correctness of the results. To get the country we then
again used geopy. The resulting data containing paper identifier, the
authors email, its associated domain, order in which the author was
cited, the institution, country, longitude and latitude were saved to a
csv file. We used the paper identifier and author order to associate
this data with the metadata to link information about authors with the
papers they authored.

@@

To get insight into collaborations within the community we built graphs
based on co-authorship of papers. To build the graph structure and
visualize it the python package \emph{networkx} was used. When
constructing the graph, we associated qualitative features --such as the
name and institution of the author that is represented by a node-- to
the nodes and edges of the graph. We also associated the number of
collaborations between different parties to the graphs as the weight of
the edge.

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_46_0.png}\end{center}
        \caption{One of the connected components of the multi-year co-authorship graph}
        \label{fig_plot}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_48_0.png}\end{center}
        \caption{Largest Component of the multiple collaboration co-authorship graph}
        \label{author_collab_large}
    \end{figure}

    Focusing in on the authors is not the only way to interpret the
co-authorship graph. One should also consider the overall structure of
the graph. As the co-authorship graph is not connected, and contains one
large main component with 1141 authors and 208 very small components
with less than 25 authors in them we focused on analyzing the structure
of the main component of the graph. First we aimed to detect
sub-communities within the graph. Modularity is one measure used to find
such communities. It measures the connection strength within a community
relative to the strength of outgoing connections. Networks with high
modularity have dense connections between the nodes within the
subcommunites but sparse connections between nodes in different
subcommunites. A non-parametric algorithm for community detection is the
Louvain Method \cite{blondel2008fast}. We used the implementation of
Thomas Aynaud \cite{louvain}. This gives a good partitioning of the
graph into 26 communities of median size 42.84. This is quite a large
number of communities, so we considered an alternative way to yield
bigger partitions. For this we looked at normalized cuts, which
partition the nodes into groups with many within-group connections and
relatively few between-group connections. The advantage of this method
is that we can choose how many cuts to make, and we are thus able to
define the size of subcommunities. Finding normalized cuts in a graph
can be done using \emph{sklearns} spectral clustering algorithm by
passing it the adjacency matrix of the graph. With this method we were
able to neatly separate the main component graph in three large
communities, as can be seen in figure \ref{normalized}

@@

To get a better overview of how co-authorship happens across countries
and institutions, we built graphs where every node represented a
country, respectively an institution. Two countries or institutions were
linked if two authors from these two institutions collaborated on a
paper together. The country graph is simple enough that it does not have
to be pruned. The institution graph on the other hand is again not
comprehensible without further processing. Thus, similar to the first
approach on pruning presented, we only considered institutions who
collaborated in at least two conferences. The main component of this
graph is shown in figure \ref{unigraph}.

@@

Graphs were visualised with the built in plot functions and graph
layouts of networkx. The force directed layout was used as it clusters
together nodes with strong connections. As this layout is produced by
iterations and based on random initial conditions, we set a fix random
seed to be able to reproduce the positioning. The size of each node was
changed based on the position of the node in the graph. Central nodes
with a lot of connections should be larger, while nodes with fewer
connections should be smaller. To achieve this we used the measures of
degree centrality for graphs with few nodes and edges and betweenness
centrality for larger graphs. Degree centrality measures the fraction of
degrees (number of connections) a node has with respect to the total
number of connections in the graph. Betweenness centrality of a node is
the sum of the fraction of all-pairs shortest paths that pass through it
\cite{borgatti2014social}. We found that degree centrality is more
appropriate for smaller graphs as their structure is still easy to
interpret. As by definition betweenness centrality takes into
consideration the degrees of a nodes neighboring nodes, it better
structures the nodes large graphs into central and less central nodes.
To decide which nodes to label, we use the same mesures by selecting the
nodes with the largest centrality, only plotting nodes who's cenrality
is in the 90th percentile of the graph. The width of the edge is based
on the weight of the edge, which in turn reflects the number of
collaboration between two nodes.

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_59_0.png}\end{center}
        \caption{Number of Contributiors by country, excluding the US}
        \label{figrep}
    \end{figure}

@@

\textbf{International Collaboration:} The U.S. again stands out as a
major player, with most collaborations happening between the U.S. and
other countries. Out of 851 papers, 153 were international
collaborations. Out of these 153, 108 involved at least one U.S.
participant. Excluding US participation we found the collaboration graph
presented in figure \ref{figcollab}. We note that some countries, such
as Japan, are not present in this graph. This means that all
international collaboration of these countries, in this case japanese
authors, also involves American authors. Additionally we observed that
many Asian nations seem to collaborate almost exclusively with each
other or Canada.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_61_0.png}\end{center}
        \caption{Collaboration between countries excluding the US}
        \label{figcollab}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_64_0.png}\end{center}
        \caption{Subcommunities by participation in CSCL. Blue: no participation in main components of CSCL. Red and orange each mark participation in one of the two main components of CSCL}
        \label{CSCL_components}
    \end{figure}

@@

The fact that normalized graph cuts will split the community into three
prompted further investigation. As will be discussed in the next
section, the graph of authors that participanted in CSCL yields two
large connected components. By coloring the authors contained in the two
CSCL components with different colors, and authors in neither with a
third color, we found a similar partitioning of the graph, as can be
observed in figure \ref{CSCL_components}. This is somewhat trivial, as
the graph is built based on collaboration in these conferences. However,
the analysis of nationalities present in the graph yielded an
interesting results. All Asian participants from nations other than
Singapore are contained in the same cluster. We also observed that one
of the CSCL clusters is closer to the group of ICLS-only participants
than the other. Additionaly, through normalized splits, we found that
the section of the normalized cut containing ICLS only authors is
overwhelmingly American.

Considering the wordclouds in figure \ref{c1}, \ref{c2}, \ref{c3} of the
most common keywords in each of the large splits, we found a first
indication that the splits also reflect some thematic differences in
research.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_66_0.png}\end{center}
        \caption{Keywords of the blue community}
        \label{c1}
    \end{figure}

@@

In our analysis, we could find many structural differences between the
two conferences. Firstly, ICLS is much larger and american dominated
than CSCL. We also observed differences in the most cited papers and the
average number of authors of a paper. By constructing the graph based on
collaboration of authors in each conference, we found that while ICLS
contains one giant component, CSCL contains two large main components.
All these results are however to be taken with a grain of salt, as for
each conference we only have data for two years, and the year 2018 is a
marked outlier. Additionaly, we are unable to find a strong difference
between the metrics of the graphs, that is diameter, density, clustering
coefficient and average path length.

@@

Knowing the affiliation of participants to an institution in a given
year, we can now find members that change institutions over the years.
Over the years of recording we found 14 participants that have changed
countries and 47 that have moved institutions. We can see that most
movement occured within the US respectively Europe, see figure
\ref{fig_move}.

@@

We observe a growth of the number of participants over the years, and by
building co-authorship graphs for each year of the conference, we
observed that the density and clustering coefficient of the graph
indicate a steady decrease, while the diameter and average shortest path
length tend to increase. This hints that the community is attracting new
member that do not collaborate with the established group of
participants. But, as the years 2018 shows a massive increase of
participation, in number of participants and number of institutions
represented, as well as a large decrease in density of the graph --which
is reduced by more than half from the previous ICLS conference-- it is
impossible to draw any statisically significant conclusion.

@@

    Since the number of documents is increasing over the years, we also
wanted to investigate more if the relevance of all the documents
published is increasing aswell. Using the Pagerank algorithm and the
In-degree of each nodes to get the most influencial documents referenced
of the exploratory citations graph that we constructed before. We
filtered the first 100 documents and then contructed a histogram of the
counts by year. We obtained the figure \ref{fig7}:

@@

    We also noticed from the first graph that some documents were more
likely to be cited together. We can say that the more two documents are
likely to be cited together the more it is probable that both documents
are about the same subject. From the table \ref{fig8} we can also notice
that some papers are more likely to cite distinct publications of the
same author, which is logic in a sense, because if some authors always
treat about the same subject then they are experts in that field and a
lot of their publications are relevant and can be cited together.

@@

    Using this graph, we also wanted to extract if there are some clusters
of papers cited together. As we explained before, we run strongly
connected algorithm to check if some cited documents are more likely to
be cited together, and this will give us clusters of cited documents
that are more likely to be about the same subjects. We can visualise one
cluster, the most important one in figure \ref{fig12}.

@@

In summary, we were able to show how the dominant some institutions and
countries are within the conference. Collaboration patterns between
countries, instututions and authors were extracted and analyzed.
Subcommunities found within the network of co-authorship reflect the
division between conferences and within the conference. Using the
analysis of the affiliation over the years we followed the migration of
participants accross institutions. Natural language preprocessing and
metadata parsing enabled the extraction of trends of keywords over the
years

Overall, due to lack of data, we were often unable to find non trivial
results; this lack of data is aggravated by the fact that the conference
of 2018 presents as an outlier in the dataset. Additionally, taking
document clustering as an example, network analysis of references were
more accurate than NLP regarding documents and references clustering.
